---
title: "P8105 Homework 1"
author: Jennifer Mizhquiri Barbecho
date: 2021-09-29
output: 
  github_document

---

# Problem 1

Below I create a data frame called **problem1_df** comprised of the following:

- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r, _libraryopen, message = FALSE}
library(tidyverse)
```


```{r _problem1_df_create}
set.seed(10)

problem1_df = tibble(
  samp = rnorm(10),
  greater_than_0_vec_logical = (samp>0),
  vec_char = c("This" , "is" , "a" , "ten", "character" , "vector" , "for" , "this" ,  "problem" , "set"),
  vec_factor = factor(c("level_a" , "level_b" , "level_c" , "level_b" , "level_c" , "level_c" , "level_a" , "level_b" , "level_c" , "level_a"))
  )

problem1_df

```
Above, **problem_df** is printed in its entirety. 

I inspect this data frame for the mean of each of the 4 variables.

```{r_mean, eval=FALSE}
mean(problem1_df$samp) ## this returns a value

mean(problem1_df$greater_than_0_vec_logical) ## this returns a value

mean(problem1_df$vec_char) ## this returns NA as the character vector is not numeric or logical

mean(problem1_df$vec_factor) ## this returns NA as the factor vector is not numeric or logical
```

As means can only be calculated from numeric or logical vectors, results are returned only for variables, _samp_ and _greater_than_0_vec_logical_. I can convert the other two variables into a numeric vector. 

```{r_numeric_conversion}
mean(as.numeric(problem1_df$vec_char)) ## returns NAs introduced by coercion

mean(as.numeric(problem1_df$vec_factor)) ## returns a value
```
 
I am able to obtain a mean value for the factor vector but not for the character vector. This makes sense as the _factor vector_ is based on levels, such that each can be ascribed a value; it should be noted that each level has not been explicitly assigned an order so the value itself may not yet be meaningful. Further, the _character vector_ does not have a value but strings of text, and we can expect that there would be no numeric mean.


# Problem 2

I will now open a dataset called **penguins**.
```{r}

data("penguins", package = "palmerpenguins")
library(tidyverse)

head(penguins)
```


Above, an excerpt of **penguins** is displayed. 

*About the data set*

This dataset provides biometric, physical location, and year data for three sets of penguins across three different islands. There are 8 variables: `r names(penguins)`. Further, the number of columns in **penguins** is `r ncol(penguins)` and the number of rows is `r nrow(penguins)`. 


*Finding the mean and other summary statistics*

You can use a summary function like summary() to get summary statistics of  **penguins**. Executing a summary command would provide insight like the mean and median of continuous variables like bill depth and bill length or the frequency by sex or island.

You can also use the mean function to find that the the mean of the flipper length variable is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm. 



*Visualizing the data*

The data in **penguins** also offers the opportunity to create a scatterplot *(fig. 1)* of bill length and flipper length by species. 
```{r _scatter, warning = FALSE}
ggplot(penguins, aes(y = flipper_length_mm, x = bill_length_mm, color = species)) + geom_point()
```

*fig 1.* _This scatterplot demonstrates that there is a weakly positive linear relationship between bill length and flipper length, regardless of penguin species. Generally, of the three species, the Adelie species has shorter bills and flippers and the Gentoo flippers has longer bills and flippers._

```{r _scatter_export, message = FALSE, warning = FALSE}
ggsave("penguins_scatter.pdf")
```
A PDF of this scatterplot (**penguins_scatter.pdf**) may also be located in this project directory. 

```


