---
title: "P8105 Homework 1"
author: Jennifer Mizhquiri Barbecho
date: 2021-09-29
output: 
  github_document

---

# Problem 1.A

Below I create a data frame called **problem1_df** comprised of the following:

- a random sample of size 10 from a standard Normal distribution
- a logical vector indicating whether elements of the sample are greater than 0
- a character vector of length 10
- a factor vector of length 10, with 3 different factor “levels”

```{r}
library(tidyverse)

set.seed(10)

problem1_df = tibble(
  samp = rnorm(10),
  greater_than_0_logical = (samp>0),
  vec_char = c("This" , "is" , "a" , "ten", "character" , "vector" , "for" , "this" ,  "problem" , "set"),
  vec_factor = factor(c("level_a" , "level_b" , "level_c" , "level_b" , "level_c" , "level_c" , "level_a" , "level_b" , "level_c" , "level_a"))
  )

problem1_df

```

I inspect the data for the mean of each of the 4 variables

```{r_mean, eval=FALSE}
mean(problem1_df$samp)#this returns a value

mean(problem1_df$greater_than_0_logical)#this returns a value

mean(problem1_df$vec_char)#this returns NA as the character vector is not numeric or logical

mean(problem1_df$vec_factor)#this returns NA as the factor vector is not numeric or logical
```

As means can only be calculated from numeric or logical vectors, results are returned only for variables, _samp_ and _greater_than_0_logical_. I can convert the other two variables into a numeric vector. 

```{r_numeric_conversion}
mean(as.numeric(problem1_df$vec_char))#returns NAs introduced by coercion

mean(as.numeric(problem1_df$vec_factor))#returns a value
```
 
I am able to obtain a mean value for the factor vector but not for the character vector. This makes sense as the _factor vector_ is based on levels, such that each can be ascribed a value; it should be noted that each level has not been assigned an order so the value itself may not yet be meaningful. Further, the _character vector_ does not have a value but strings of text, and it makes sense that it would not have a numeric mean. 


# Problem 2

I will now move
```{r}

data("penguins", package = "palmerpenguins")
library(tidyverse)

view(penguins)
names(penguins)
summary(penguins)

nrow(penguins)
ncol(penguins)

pull(penguins)

penguins$flipper_length_mm
class(penguins$flipper_length_mm)

mean(penguins$flipper_length_mm, na.rm = TRUE)
```
